---
title: "Pan-Arctic DSL - Arctic current velocities and directions"
author: "Pierre Priou - pierre.priou@mi.mun.ca"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(dtplyr)
library(lubridate)
library(plotly)
library(cowplot)
library(raster)
library(sf)
library(metR)
library(cmocean)
library(tidync)
```

```{r figure-theme}
theme_set(theme_bw())
theme_update(panel.border = element_blank(), 
             axis.line = element_line(),
             axis.text = element_text(size = 10),
             axis.title = element_text(size = 11),
             legend.title = element_text(size = 11),
             strip.text.x = element_text(size = 11, face = "plain", 
                                         hjust = 0.5),
             strip.background = element_rect(colour = "transparent", 
                                             fill = "transparent"),
             legend.margin = margin(0, 0, 0, 0),
             legend.box.margin = margin(0, 0, -8, 0),
             plot.margin = unit(c(0.075, 0.075, 0.075, 0.075), "in"))
```

# Aim

This document creates maps of annual mean current velocity and directions in 2015, 2016, and 2017. 

# NetCDF to csv

First I convert Arctic Ocean Physics Reanalysis data from NetCDF to csv files. 

```{r NetCDF-csv, eval=FALSE}
# Empty dataframe
velo <- data.frame() 

# List files per year
file_list <- list.files("data/ocean_currents/nc", pattern = "*.nc", full.names = T) 

# Loop through each file
for (i in file_list) { 
  
  # Temperature, salinity, depth, and current velocity
  velo_tmp <- tidync(i) %>%
    activate("D0,D1,D3,D2") %>% 
    hyper_tibble()
  
  # Coordinates
  coord_tmp <- tidync(i) %>%
    activate("D0,D1") %>% 
    hyper_tibble() %>%
    # Find extent of grid
    mutate(xmin = min(x),
           xmax = max(x),
           ymin = min(y), 
           ymax = max(y))
  
  # Retrieve projection details
  r <- terra::rast(i)
  proj_tmp <- tibble(proj = crs(r, proj = T),
                     res_x = r@ptr$res[1],
                     res_y = r@ptr$res[2])

  # Combine datasets
  tmp <- left_join(velo_tmp, coord_tmp, by = c("x", "y")) %>%
    # Add projection details
    bind_cols(., proj_tmp) %>%
    # Calculate date and add extents
    mutate(date = as.POSIXct(time * 3600, origin = "1950-01-01 00:00:00", tz = "UTC")) %>%
    # Rename variables
    rename(lat = latitude, 
           lon = longitude) %>%
    # Only keep velocity data 
    dplyr::select(date, lat, lon, x, y, depth, vxo, vyo, proj, res_x, res_y,
                  xmin, xmax, ymin, ymax)
  
  # Append data in a single file
  velo <- bind_rows(velo, tmp)
  print(paste0("Processing of ", i, " over."))
}

# Write csv file
write_csv(velo, file = "data/ocean_currents/velocity_318m_CMEMS.csv")

rm(velo, file_list, i, velo_tmp, coord_tmp, r, proj_tmp, tmp)
```

# Data tidying

I load the csv file and tidy it. 

```{r load-data}
# List files
velo_raw <- read_csv("data/ocean_currents/velocity_318m_CMEMS.csv", show_col_types = F)

# Get projection, cell resolution, and extents of data
proj_current <- velo_raw$proj[1]
res_current <- c(velo_raw$res_x[1], velo_raw$res_x[1])
extent_current <- c(velo_raw$xmin[1], velo_raw$xmax[1], velo_raw$ymin[1], velo_raw$ymax[1])

# Remove projection variables from tibble
velo <- velo_raw %>%
  dplyr::select(-proj, -res_x, -res_y) %>%
  # Add month and year, and calculate velocity (in m/s)
  mutate(year = year(date),
         month = month(date),
         velocity = sqrt(vxo^2 + vyo^2))
```

The original projection of the data is: *`r proj_current`*

Each pixel represents a `r paste0(res_current[1] * 100, " km x ", res_current[2] * 100, " km")` cell.

The variables are:

-   time: time in hours since 1950-01-01 00:00:00
-   latitude: latitude
-   longitude: longitude
-   x: x-coordinate (polar stereographic x coordinate)
-   y: y-coordinate (polar stereographic y coordinate)
-   vxo: Zonal velocity m/s (x)
-   vyo: Meridional velocity m/s (y)

```{r load-coastline-data, message=FALSE, warning=FALSE, include=FALSE}
proj_coast <- raster(extent(extent_current), crs = proj_current, resolution = res_current)

# Coastline shapefile
coast <- rgdal::readOGR("data/bathy/ne_10m_land.shp", verbose = F) %>% 
  # spTransform(CRSobj = crs("EPSG:4326")) %>% 
  crop(extent(-180, 180, 0, 90)) %>%
  # Project shapefile in new proj
  spTransform(CRSobj = crs(proj_current)) %>% 
  # Convert to a dataframe for ggplot
  fortify() %>% 
  rename(x = long, y = lat) %>%
  mutate(x = x * 10^-5,
         y = y * 10^-5)

rm(proj_coast)
```

Here is a plot of current velocity at 318 m depth in October 2016.

```{r plot-velocity-oct-2016-12.5km, fig.height=8.5, fig.width=10}
velo %>%
  filter(year == 2016 & month == 10) %>%
  ggplot(aes(x = x, y = y)) +
  geom_tile(aes(fill = velocity), alpha = 0.8) +
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  geom_vector(aes(dx = vxo, dy = vyo, mag = velocity), 
              arrow.angle = 30, arrow.type = "open", arrow.length = 1,
              pivot = 0, preserve.dir = TRUE, direction = "ccw") +
  scale_color_viridis_c(option = "turbo", 
                        limits = c(0, 0.075), oob = scales::squish) + 
  scale_fill_cmocean(name = "speed") +
  coord_fixed(xlim = c(-27, 27), ylim = c(-35, 18), expand = F) +
  guides(mag = "none", alpha = "none") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
```

# Reduce data resolution

Define projections. 

```{r current-projection}
# Set current projection (12.5 x 12.5 km)
r_current <- raster(extent(extent_current), crs = proj_current, resolution = res_current)

# Set new projection 
res_new <- c(1, 1)
r_new <- raster(extent(extent_current), crs = proj_current, resolution = res_new)
```

Calculate the mean velocity and direction for each grid cell per year.

```{r rasterize-data}
# Empty dataframe
velo_new <- data.frame()

# Loop through each year
for (i in seq(2015, 2017, 1)) {
  
  # Select data of that year
  data_tmp <- velo %>%
    filter(year == i)
  
  # Convert to spatial data
  velo_tmp <- SpatialPointsDataFrame(
    SpatialPoints(cbind(data_tmp$x, data_tmp$y), proj4string = crs(r_current)),
    data.frame(lat = data_tmp$lat,
               lon = data_tmp$lon,
               vxo = data_tmp$vxo,
               vyo = data_tmp$vyo)) %>%
    # Rasterize on the 100 x 100 km cells
    rasterize(., r_new, fun = mean, na.rm = T) %>% 
    # Remove ID layer
    dropLayer(1) %>% 
    # Convert raster to data frame
    rasterToPoints() %>% 
    as.data.frame() %>% 
    # Add year
    mutate(year = i, 
           velocity = sqrt(vxo^2 + vyo^2))
  
  # Append into a single dataframe
  velo_new <- bind_rows(velo_new, velo_tmp)
}

# Remove unused variables
rm(i, data_tmp, velo_tmp)
```

The following plot show annual mean current velocity in 2016 with a `r paste0(res_new[1] * 100, " km x ", res_new[2] * 100, " km")` cell resolution.

```{r plot-velocity-oct-2016-new-res, fig.height=8.5, fig.width=10}
velo_new %>%
  filter(year == 2016) %>%
  ggplot(aes(x = x, y = y)) +
  geom_tile(aes(fill = velocity), alpha = 0.8) +
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  geom_vector(aes(dx = vxo, dy = vyo, mag = velocity), 
              arrow.angle = 30, arrow.type = "open", arrow.length = 1,
              pivot = 0, preserve.dir = TRUE, direction = "ccw") +
  scale_color_viridis_c(option = "turbo", 
                        limits = c(0, 0.075), oob = scales::squish) + 
  scale_fill_cmocean(name = "speed") +
  coord_fixed(xlim = c(-27, 27), ylim = c(-35, 18), expand = F) +
  guides(mag = "none", alpha = "none") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
```
