---
title: "PanArctic DSL - Remote sensing"
author: "[Pierre Priou](mailto:pierre.priou@mi.mun.ca)"
date: "`r format(Sys.time(), '%Y/%m/%d at %H:%M')`"
output: github_document
always_allow_html: true
---

# Package loading

```{r package-loading, message=FALSE}
library(tidyverse)  # Tidy code
library(lubridate)  # Deal with dates
library(tidync)     # Read NetCDF
library(raster)     # Rasterize data
library(ecmwfr)     # Download Copernicus data
```

# Sea ice data

Code that downloads sea ice data from the [Copernicus website]( https://cds.climate.copernicus.eu/cdsapp#!/dataset/satellite-sea-ice-concentration?tab=overview). We use the daily sea ice concentration derived from satellite observations from ECMWF. Data is downloaded in two batches because ECMWF changed their Climate Data Record (CDR) on 2016-01-01. I changed the sea ice data projection from Lambert Azimuthal Equal Area (EPSG:6931) to WGS84 (EPSG:4326) and match the resolution to that of backscatter anomalies (2 deg longitude * 1 deg latitude).

## CDR data 2014-01-01 - 2015-12-31

```{r download-and-reproject-seaice-CDR, message=FALSE}
# Define projections: latlon (acoustic and CTD data), and laea (sea ice data)
arctic_latlon <- raster(extent(-180,180,0,90), crs = "+init=epsg:4326", res = c(2, 1))
arctic_laea <- raster(extent(-5400,5400,-5400,5400), crs = "+init=epsg:6931", res = c(25, 25))
# Date range of interest
dates <- seq(ymd("2014-01-01"), ymd("2015-12-31"), 1)
# Download data
for (d in dates) {
  # Format years, month, and day to be correctly read in the request
  year = as.character(year(as.Date(d, origin = "1970-01-01")))
  month = if_else(as.numeric(month(as.Date(d, origin =" 1970-01-01"))) < 10,
                  paste0("0", as.character(month(as.Date(d, origin = "1970-01-01")))),
                  as.character(month(as.Date(d, origin = "1970-01-01"))))
  day = if_else(as.numeric(day(as.Date(d, origin = "1970-01-01"))) < 10,
                paste0("0", as.character(day(as.Date(d, origin = "1970-01-01")))),
                as.character(day(as.Date(d, origin = "1970-01-01"))))
  request <- list(version = "v2", # Create request
                  variable = "all",
                  format = "zip",
                  origin = "eumetsat_osi_saf",
                  region = "northern_hemisphere",
                  cdr_type = "cdr",
                  year = year,
                  month = month,
                  day = day,
                  dataset_short_name = "satellite-sea-ice-concentration",
                  target = "download.zip")
  print(paste0("Downloading and tidying ", request$dataset_short_name, " from ", year, "-", month, "-", day))
  tmp_raw <- wf_request(user = "113650", # Download requested file
                        request = request, 
                        transfer = T, 
                        verbose = F,
                        path = "data/remote_sensing/sea_ice/tmp") %>%
    unzip(exdir = "data/remote_sensing/sea_ice/tmp") %>%  # Unzip file
    tidync() # Read sea ice data
  # tmp_raw <- tidync("data/remote_sensing/ice_conc_nh_ease2-250_icdr-v2p0_201703091200.nc")
  tmp_coord <- tmp_raw %>% # Extract latlon and laea coordinates
    activate("D2,D3") %>%
    hyper_tibble()
  tmp_seaice <- tmp_raw %>%
    activate("D2,D3,D0") %>%
    hyper_tibble() %>%
    left_join(., tmp_coord, by = c("xc", "yc")) %>% # Join latlon coordinates to sea ice data
    mutate(date = ymd(format(as_datetime(time, origin = "1978-01-01", tz = "UTC"), format = "%Y%m%d"))) %>%
    filter(status_flag != c(1, 2, 16)) # Remove data from land (1), lakes (2), and possible dalse ice (16)
  tmp <- SpatialPointsDataFrame(SpatialPoints(cbind(tmp_seaice$lon, tmp_seaice$lat), # transform in spatial point data frame
                                              proj4string = CRS("+init=epsg:4326")),
                                data.frame(date = as.numeric(tmp_seaice$date), 
                                           ice_conc = tmp_seaice$ice_conc, 
                                           total_se = tmp_seaice$total_standard_error,
                                           smearing_se = tmp_seaice$smearing_standard_error,
                                           algorithm_se = tmp_seaice$algorithm_standard_error)) %>%
    rasterize(., arctic_latlon, fun = mean, na.rm = T) %>% # Rasterize data in latlon
    dropLayer(1) %>% # Remove ID layer
    crop(., y = extent(-155, 35, 66, 85)) %>% # Crop raster to areas of interest
    rasterToPoints() %>% # Convert raster to data frame
    as.data.frame() %>%
    rename(lon = x, lat = y) %>% # Rename variables
    mutate(date = as_date(date)) # Reformat date 
  write_csv(tmp, file = paste0("data/remote_sensing/sea_ice/", year, month, day, "_ice_conc.csv")) # Save data as a csv file
  file.remove(list.files("data/remote_sensing/sea_ice/tmp", full.names = T)) # Delete temporary netcdf and zip files
}
rm(request, tmp_raw, tmp_coord, tmp_seaice, tmp) # Remove unused variables
```

## ICDR data 2016-01-01 - 2017-12-31

```{r download-and-reproject-seaice-ICDR, message=FALSE}
# Define projections: latlon (acoustic and CTD data), and laea (sea ice data)
arctic_latlon <- raster(extent(-180,180,0,90), crs = "+init=epsg:4326", res = c(2, 1))
arctic_laea <- raster(extent(-5400,5400,-5400,5400), crs = "+init=epsg:6931", res = c(25, 25))
# Date range of interest
dates <- seq(ymd("2016-01-01"), ymd("2017-12-31"), 1)
# Download data
for (d in dates) {
  # Format years, month, and day to be correctly read in the request
  year = as.character(year(as.Date(d, origin = "1970-01-01")))
  month = if_else(as.numeric(month(as.Date(d, origin =" 1970-01-01"))) < 10,
                  paste0("0", as.character(month(as.Date(d, origin = "1970-01-01")))),
                  as.character(month(as.Date(d, origin = "1970-01-01"))))
  day = if_else(as.numeric(day(as.Date(d, origin = "1970-01-01"))) < 10,
                paste0("0", as.character(day(as.Date(d, origin = "1970-01-01")))),
                as.character(day(as.Date(d, origin = "1970-01-01"))))
  request <- list(version = "v2", # Create request
                  variable = "all",
                  format = "zip",
                  origin = "eumetsat_osi_saf",
                  region = "northern_hemisphere",
                  cdr_type = "icdr",
                  year = year,
                  month = month,
                  day = day,
                  dataset_short_name = "satellite-sea-ice-concentration",
                  target = "download.zip")
  print(paste0("Downloading and tidying ", request$dataset_short_name, " from ", year, "-", month, "-", day))
  tmp_raw <- wf_request(user = "113650", # Download requested file
                        request = request, 
                        transfer = T, 
                        verbose = F,
                        path = "data/remote_sensing/sea_ice/tmp") %>%
    unzip(exdir = "data/remote_sensing/sea_ice/tmp") %>%  # Unzip file
    tidync() # Read sea ice data
  # tmp_raw <- tidync("data/remote_sensing/ice_conc_nh_ease2-250_icdr-v2p0_201703091200.nc")
  tmp_coord <- tmp_raw %>% # Extract latlon and laea coordinates
    activate("D2,D3") %>%
    hyper_tibble()
  tmp_seaice <- tmp_raw %>%
    activate("D2,D3,D0") %>%
    hyper_tibble() %>%
    left_join(., tmp_coord, by = c("xc", "yc")) %>% # Join latlon coordinates to sea ice data
    mutate(date = ymd(format(as_datetime(time, origin = "1978-01-01", tz = "UTC"), format = "%Y%m%d"))) %>%
    filter(status_flag != c(1, 2, 16)) # Remove data from land (1), lakes (2), and possible dalse ice (16)
  tmp <- SpatialPointsDataFrame(SpatialPoints(cbind(tmp_seaice$lon, tmp_seaice$lat), # transform in spatial point data frame
                                              proj4string = CRS("+init=epsg:4326")),
                                data.frame(date = as.numeric(tmp_seaice$date), 
                                           ice_conc = tmp_seaice$ice_conc, 
                                           total_se = tmp_seaice$total_standard_error,
                                           smearing_se = tmp_seaice$smearing_standard_error,
                                           algorithm_se = tmp_seaice$algorithm_standard_error)) %>%
    rasterize(., arctic_latlon, fun = mean, na.rm = T) %>% # Rasterize data in latlon
    dropLayer(1) %>% # Remove ID layer
    crop(., y = extent(-155, 35, 66, 85)) %>% # Crop raster to areas of interest
    rasterToPoints() %>% # Convert raster to data frame
    as.data.frame() %>%
    rename(lon = x, lat = y) %>% # Rename variables
    mutate(date = as_date(date)) # Reformat date 
  write_csv(tmp, file = paste0("data/remote_sensing/sea_ice/", year, month, day, "_ice_conc.csv")) # Save data as a csv file
  file.remove(list.files("data/remote_sensing/sea_ice/tmp", full.names = T)) # Delete temporary netcdf and zip files
}
rm(request, tmp_raw, tmp_coord, tmp_seaice, tmp) # Remove unused variables
```

Plot data to see if reprojection worked.

```{r plot-seaice-latlon-cdr, message=FALSE, fig.align='center', fig.height=3.25, fig.width=6, dpi=600}
read_csv("data/remote_sensing/sea_ice/20150212_ice_conc.csv") %>%
  ggplot(aes(x = lon, y = lat, fill = ice_conc)) + 
  geom_tile() + 
  scale_fill_viridis_c() +
  ggtitle("2015-02-12 Ice concentration (%) CDR")
```

```{r plot-seaice-latlon-icdr, message=FALSE, fig.align='center', fig.height=3.25, fig.width=6, dpi=600}
read_csv("data/remote_sensing/sea_ice/20170212_ice_conc.csv") %>%
  ggplot(aes(x = lon, y = lat, fill = ice_conc)) + 
  geom_tile() + 
  scale_fill_viridis_c() +
  ggtitle("2017-02-12 Ice concentration (%) ICDR")
```
