---
title: "PanArctic DSL - Analyses"
author: "[Pierre Priou](mailto:pierre.priou@mi.mun.ca)"
date: "`r format(Sys.time(), '%Y/%m/%d at %H:%M')`"
output: github_document
always_allow_html: true
---

# Package and data loading

```{r package-loading, message=FALSE}
# Load packages
library(tidyverse)              # Tidy code
library(lubridate)              # Deal with dates
library(kableExtra)             # Pretty tables
library(cowplot)                # Plots on a grid
library(rgdal)                  # Read shapefiles
library(rworldxtra)             # Higher resolution coastline data
library(marmap)                 # Bathymetry data of the Arctic Ocean
library(ggforce)                # Draw polygons
library(cmocean)                # Pretty color palettes
source("R/getNOAA.ice.bathy.R") # Load bathy data from NOAA
# Custom figure theme
theme_set(theme_bw())
theme_update(axis.text = element_text(size = 9),
             axis.title = element_text(size = 9),
             strip.text.x = element_text(size = 9, face = "plain", hjust = 0.5),
             strip.background = element_rect(colour = "transparent", fill = "transparent"),
             legend.title = element_text(size = 9),
             legend.margin = margin(0, 0, 0, 0),
             legend.box.margin = margin(0, 0, -8, 0),
             panel.grid = element_blank(), 
             plot.margin = unit(c(0, 0, 0, 0), "in"))
options(dplyr.summarise.inform = F) # Suppress summarise() warning
```

```{r data-loading, message=FALSE}
# Geographical data
coastlines_110m <- readOGR("data/bathy/ne_110m_land.shp", verbose = F) %>% 
  fortify() %>%
  rename(lon = long, region = id)
coastlines_50m <- readOGR("data/bathy/ne_50m_land.shp", verbose = F) %>% 
  fortify() %>%
  rename(lon = long, region = id)
coastlines_10m <- readOGR("data/bathy/ne_10m_land.shp", verbose = F) %>% 
  fortify() %>%
  rename(lon = long, region = id)
# Bathy data
bathy_df <- getNOAA.ice.bathy(lon1 = -180, lon2 = 180, lat1 = 50, lat2 = 90,
                              resolution = 6, keep = T, path = "data/bathy/") %>%
  fortify.bathy() %>%
  rename(lon = x, lat = y, depth = z) %>%
  mutate(depth_d=factor(case_when(between(depth, -50, Inf) ~ "0-50",
                                  between(depth, -100, -50) ~ "50-100",
                                  between(depth, -200, -100) ~ "100-200",
                                  between(depth, -500, -200) ~ "200-500",
                                  between(depth, -1000, -500) ~ "500-1000",
                                  between(depth, -2000, -1000) ~ "1000-2000",
                                  between(depth, -4000, -2000) ~ "2000-4000",
                                  between(depth, -Inf, -4000) ~ "4000-8000"),
                        levels=c(">0","0-50","50-100","100-200","200-500","500-1000","1000-2000","2000-4000","4000-8000")))
# Bathy data
bathyHD_df <- getNOAA.ice.bathy(lon1 = -180, lon2 = 180, lat1 = 50, lat2 = 90,
                              resolution = 2, keep = T, path = "data/bathy/") %>%
  fortify.bathy() %>%
  rename(lon = x, lat = y, depth = z) %>%
  mutate(depth_d=factor(case_when(between(depth, -50, Inf) ~ "0-50",
                                  between(depth, -100, -50) ~ "50-100",
                                  between(depth, -200, -100) ~ "100-200",
                                  between(depth, -500, -200) ~ "200-500",
                                  between(depth, -1000, -500) ~ "500-1000",
                                  between(depth, -2000, -1000) ~ "1000-2000",
                                  between(depth, -4000, -2000) ~ "2000-4000",
                                  between(depth, -Inf, -4000) ~ "4000-8000"),
                        levels=c(">0","0-50","50-100","100-200","200-500","500-1000","1000-2000","2000-4000","4000-8000")))
# Areas of interest
area_BF_CAA <- data.frame(x = c(-153, -141.5, -93, -94, -153), y = c(75, 69, 69, 78, 78))
area_BB <- data.frame(x = c(-85, -65, -51.5, -51.5, -85), y = c(72, 66.3, 66.3, 82, 82))
area_SV <- data.frame(x = c(40, 21, 0, 0, 40), y = c(80, 78, 77, 85, 85))
area_BF_CAA_mercator <- data.frame(x = c(-153.5, -153.5, -90, -90), y = c(67, 78, 78, 67))
area_BB_mercator <- data.frame(x = c(-86, -65, -50, -50, -86), y = c(72, 66, 66, 82.5, 82.5))
area_SV_mercator <- data.frame(x = c(37, 37, 2, 2), y = c(75, 85, 85, 75))
area_BF_CAA_2 <- data.frame(x = c(-158, -158, -93, -93, -158), y = c(79, 68.5, 68.5, 79, 79))
area_BB_2 <- data.frame(x = c(-86, -65, -51.5, -51.5, -86), y = c(72, 65.5, 65.5, 83, 83))
area_SV_2 <- data.frame(x = c(46, 46, 0, 0, 46), y = c(85, 76, 76, 85, 85))
# Acoustic data
load("data/acoustics/MVBS_2015_2017.RData")
# CTD data
load("data/CTD/CTD_2015_2019.RData")
# Trawl data
load("data/nets/trawl_mwt_2015_2017.RData")
# Stats data
# load("data/acoustics/kruskal_dunn_spatial_interannual.RData")
# Primary production data
load("data/remote_sensing/remote_sensing_chl.RData")
```

# Plots 

## Figure 1. Bathy map with sampling locations

Careful this map takes forever to compute (> 18 h) when bathy has to be drawn.

```{r fig1-map-station, fig.align='center', fig.height=3.25, fig.width=6, dpi=600}
trawl_loc <- trawl_station %>%
  group_by(date) %>%
  summarise(lat = mean(lat),
            lon = mean(lon))
MVBS %>%
  mutate(lat = round(lat, 1),
         lon = round(lon, 1)) %>%
  group_by(lat,lon) %>%
  summarise(lat = mean(lat), 
            lon = mean(lon)) %>%
  ungroup() %>%
  ggplot(aes(x = lon, y = lat)) +
  # Plot bathy
  # geom_tile(data = bathyHD_df, aes(x = lon, y = lat, fill = depth_d)) +
  # scale_fill_cmocean("Depth (m)", name = "deep", discrete = T, na.value = NA, alpha = 0.6) +
  # Plot coastlines
  geom_polygon(data = coastlines_10m, aes(x = lon, y = lat, group = group), fill = "grey70") +
  # Arctic circle
  geom_segment(aes(x = -180, xend = 0, y = 66.5, yend = 66.5), col = "grey40", lty = 2, size = 0.3) +
  geom_segment(aes(x = 0, xend = 180, y = 66.5, yend = 66.5), col = "grey40", lty = 2, size = 0.3) +
  # Areas of interest
  geom_shape(data=area_BF_CAA_2, aes(x=x,y=y), fill=NA, col="black", lwd=0.2, lty=1) +
  geom_shape(data=area_BB_2, aes(x=x,y=y), fill=NA, col="black", lwd=0.2, lty=1) +
  geom_shape(data=area_SV_2, aes(x=x,y=y), fill=NA, col="black", lwd=0.2, lty=1) +
  # Acoustic data
  geom_point(size = 0.7, shape = 4, color = "black") +
  # Trawl locations
  geom_point(data = trawl_loc, aes(x = lon, y = lat), shape = 23, size = 1.4, color = "black", fill = "orange") +
  # Change projection
  coord_map("azequalarea", ylim = c(66, 90), xlim = c(-160, 70), orientation = c(90,0,-60)) +
  guides(fill = guide_legend(keywidth = 0.2, keyheight = 0.2, default.unit = "in"), color = "none") +
  theme(legend.position = "right", panel.border = element_rect(fill = NA), axis.text = element_blank(),
        axis.ticks = element_blank(), axis.title = element_blank())
```

## Figure 2. Vertical S~V~ profiles

Code that works and used for producing the plot in the manuscript from 2021_12_17

```{r fig2-Sv-profile-area-year, echo=F, warning=F,fig.align='center', dpi=100, fig.width=6.5, fig.height=3, dpi=600}
# Custom color palette
col_pal <- c("#80CBB1", "#347BA5", "#302346")
# Calculate mean and median vertical profile for each area
Sv_raster %>%
  # Calculate mean and median Sv profiles per area per year
  group_by(depth, area, year) %>%
  summarise(sv_lin_median_areayear=median(sv_lin_median),
            sv_lin_q2.5_areayear=median(sv_lin_q2.5),
            sv_lin_q97.5_areayear=median(sv_lin_q97.5)) %>%
  mutate(Sv_median=10*log10(sv_lin_median_areayear), 
         Sv_q2.5=10*log10(sv_lin_q2.5_areayear),
         Sv_q97.5=10*log10(sv_lin_q97.5_areayear)) %>%
  ungroup() %>%
  # Plotting details
  # Remove empty water column and very low backscatter
  mutate(Sv_median_1=if_else(Sv_median < -200, NaN, Sv_median),
         Sv_q2.5_1=if_else(Sv_q2.5 < -200, NaN, Sv_q2.5),
         Sv_q97.5_1=if_else(Sv_q97.5 < -105, NaN, Sv_q97.5)) %>%
  # Replace low values by min limit for plotting continuous ribbon
  mutate(Sv_median_2=if_else(is.na(Sv_median_1)==F & Sv_median_1<=-105, -105, Sv_median_1),
         Sv_q2.5_2=if_else(is.na(Sv_median_1)==F & is.na(Sv_q2.5_1)==T & is.na(Sv_q97.5_1)==F, -105,
                     if_else(is.na(Sv_median_1)==T & is.na(Sv_q2.5_1)==T & is.na(Sv_q97.5_1)==F, -105, 
                     if_else(is.na(Sv_q2.5_1)==F & Sv_q2.5_1 <=-105, -105, Sv_q2.5_1))),
         Sv_q97.5_2=Sv_q97.5_1) %>%
  mutate(Sv_median_3=if_else(Sv_median_2==-105 & Sv_q2.5_2==-105 & is.na(Sv_q97.5_2)==T, NaN, Sv_median_2),
         Sv_q2.5_3=if_else(Sv_median_2==-105 & Sv_q2.5_2==-105 & is.na(Sv_q97.5_2)==T, NaN, Sv_median_3),
         Sv_q97.5_3=Sv_q97.5_2) %>%
  arrange(year, area, depth) %>%
  mutate(area=factor(if_else(area=="BF_CAA", "Beaufort Sea &\nCan. Arctic Archipelago",
                       if_else(area=="BB", "Baffin Bay",
                       if_else(area=="SV", "Svalbard","Other"))),
                       levels=c("Beaufort Sea &\nCan. Arctic Archipelago","Baffin Bay","Svalbard"))) %>%
  # Plot
  ggplot() +
  geom_vline(xintercept=200, lty=2, col="grey20") +
  geom_ribbon(aes(x=depth, ymin=Sv_q2.5_2, ymax=Sv_q97.5_2, fill=as.factor(year), group=year), alpha=0.2) +
  geom_line(aes(x=depth, y=Sv_median_2, col=as.factor(year), group=year)) +
  scale_x_reverse("Depth (m)", breaks=seq(0,1000,200)) +
  geom_hline(yintercept=-105, col="white") +
  scale_y_continuous(expression("S"[V]*" (dB re 1 m"^-1*")"), breaks=seq(-200,0,10), limits=c(-105,-65), expand=c(0,0)) +
  scale_color_manual(values=col_pal) +
  scale_fill_manual(values=col_pal) +
  coord_flip() +
  facet_grid(~area) +
  theme(legend.title=element_blank(),
        legend.position=c(0.94,0.275),
        legend.background=element_blank(),
        legend.key=element_blank(), 
        panel.border=element_blank(),
        axis.line=element_line())
```


**WORK IN PROGRESS**

I'm trying to streamline the code.

```{r Sv-profiles-area-year-in-progress, eval=FALSE, fig.align='center', fig.width=6.5, fig.height=3, dpi=600}
col_pal <- c("#80CBB1", "#347BA5", "#302346") # Custom colour palette
Sv_area_year <- Sv_raster %>%
  group_by(depth, area, year) %>%
  # Calculate mean and median Sv profiles per area per year
  summarise(sv_lin_median_areayear = median(sv_lin_median),
            sv_lin_q2.5_areayear = median(sv_lin_q2.5),
            sv_lin_q97.5_areayear = median(sv_lin_q97.5)) %>%
  # Transform into volume backscattering strength (dB)
  mutate(Sv_median = 10 * log10(sv_lin_median_areayear), 
         Sv_q2.5 = 10 * log10(sv_lin_q2.5_areayear),
         Sv_q97.5 = 10 * log10(sv_lin_q97.5_areayear)) %>%
  ungroup() %>%
  # Plotting details
  # # # Remove empty water column and very low backscatter
  # mutate(Sv_median_1 = if_else(Sv_median < -200, NaN, Sv_median),
  #        Sv_q2.5_1 = if_else(Sv_q2.5 < -200, NaN, Sv_q2.5),
  #        Sv_q97.5_1 = if_else(Sv_q97.5 < -105, NaN, Sv_q97.5)) %>%
  # Remove empty water column and very low backscatter
  mutate(Sv_q97.5_1 = if_else(Sv_q97.5 < -105, NaN, Sv_q97.5),
         Sv_q2.5_1 = case_when(Sv_q97.5 < -105 ~ NaN,
                               Sv_q2.5 < -105 ~ NaN,
                               Sv_q2.5 >= -105 ~ Sv_q2.5)) %>%
  # Replace low values by min limit for plotting continuous ribbon
  mutate(Sv_q2.5_2 = if_else(is.na(Sv_q2.5_1) == T & is.na(Sv_q97.5_1) == F, -105,
                             if_else(is.na(Sv_q2.5_1) == T & is.na(Sv_q97.5_1) == F, -105, 
                                     if_else(is.na(Sv_q2.5_1) == F & Sv_q2.5_1 <= -105, -105, Sv_q2.5_1))),
         Sv_q97.5_2 = Sv_q97.5_1) %>%
  # mutate(Sv_median_2 = if_else(is.na(Sv_median_1) == F & Sv_median_1 <= -105, -105, Sv_median_1),
  #        Sv_q2.5_2 = if_else(is.na(Sv_median_1) == F & is.na(Sv_q2.5_1) == T & is.na(Sv_q97.5_1) == F, -105,
  #                    if_else(is.na(Sv_median_1) == T & is.na(Sv_q2.5_1) == T & is.na(Sv_q97.5_1) == F, -105, 
  #                    if_else(is.na(Sv_q2.5_1) == F & Sv_q2.5_1 <= -105, -105, Sv_q2.5_1))),
  #        Sv_q97.5_2 = Sv_q97.5_1) %>%
  arrange(year, area, depth) %>%
  # filter(area == "BF_CAA") %>%
  # mutate(area=factor(case_when(area == "BF_CAA" ~ "Beaufort Sea &\nCan. Arctic Archipelago",
  #                              area == "BB" ~ "Baffin Bay",
  #                              area == "SV" ~ "Svalbard"),
  #                    levels = c("Beaufort Sea &\nCan. Arctic Archipelago", "Baffin Bay", "Svalbard"))) %>%
  # Plot
  ggplot() +
  geom_vline(xintercept = 200, lty = 2, col = "grey20") +
  # geom_ribbon(aes(x = depth, ymin = Sv_q2.5_2, ymax = Sv_q97.5_2, fill = as.factor(year), group = year), alpha = 0.2) +
  # geom_line(aes(x = depth, y = Sv_median_2, col = as.factor(year), group = year)) +
  geom_ribbon(aes(x = depth, ymin = Sv_q2.5, ymax = Sv_q97.5, fill = as.factor(year), group = year), alpha = 0.2, na.rm=T) +
  geom_line(aes(x = depth, y = Sv_median, col = as.factor(year), group = year)) +
  scale_x_reverse("Depth (m)", breaks = seq(0, 1000, 200)) +
  # geom_hline(yintercept = -105, col = "white") +
  # coord_cartesian(ylim = c(-105, -65)) +
  # scale_y_continuous(expression("S"[V]*" (dB re 1 m"^-1*")"), breaks = seq(-200, 0, 10), limits = c(-105, -65), expand = c(0, 0)) +
  scale_color_manual(values = col_pal) +
  scale_fill_manual(values = col_pal) +
  coord_flip(ylim = c(-100, -65)) +
  facet_grid( ~ area) +
  theme(legend.title = element_blank(), legend.position = c(0.94, 0.275), legend.background = element_blank(),
        legend.key = element_blank(),panel.border = element_blank(), axis.line = element_line())
Sv_area_year
```

## Figure 3. Map of mesopelagic backscatter hotspots

```{r fig3-map-sa-anomalies, fig.align='center', fig.height=3.25, fig.width=6, dpi=600}
col_pal <- c("#80cdc1",  "#f5f5f5", "#dfc27d", "#a6611a")
# Calculate mean sa anomaly over three years of data per area
map_sa_anomaly <- SA_anomaly_2deg %>%
  group_by(lat, lon, area, region) %>%
  summarise(year_sampled = n(),
            mean_ano_sa_int = mean(anomaly_NASC_int, na.rm = T)) %>%
  ungroup() %>%
  mutate(mean_ano_sa_int_d = factor(case_when(mean_ano_sa_int < -1 ~ "<-1",
                                              mean_ano_sa_int >= -1 & mean_ano_sa_int < -0.5 ~ "[-1; -0.5[",
                                              mean_ano_sa_int >= -0.5 & mean_ano_sa_int <= 0.5 ~ "[-0.5; 0.5]",
                                              mean_ano_sa_int > 0.5 & mean_ano_sa_int <= 1 ~ "]0.5; 1]",
                                              mean_ano_sa_int > 1 ~ ">1"),
                                    levels = c("<-1", "[-1; -0.5[", "[-0.5; 0.5]", "]0.5; 1]", ">1"))) %>%
  ggplot(aes(x = lon, y = lat)) +
  # Longitude lines
  geom_segment(aes(x = -180, xend = -180, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = -150, xend = -150, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = -120, xend = -120, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = -90, xend = -90, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = -60, xend = -60, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = -30, xend = -30, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 0, xend = 0, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 180, xend = 180, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 150, xend = 150, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 120, xend = 120, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 90, xend = 90, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 60, xend = 60, y = 60, yend = 90), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 30, xend = 30, y = 60, yend = 90), col = "grey90", size = 0.3) +
  # Latitude lines
  geom_segment(aes(x = -180, xend = 0, y = 60, yend = 60), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 0, xend = 180, y = 60, yend = 60), col = "grey90", size = 0.3) +
  geom_segment(aes(x = -180, xend = 0, y = 70, yend = 70), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 0, xend = 180, y = 70, yend = 70), col = "grey90", size = 0.3) +
  geom_segment(aes(x = -180, xend = 0, y = 80, yend = 80), col = "grey90", size = 0.3) +
  geom_segment(aes(x = 0, xend = 180, y = 80, yend = 80), col = "grey90", size = 0.3) +
  # Plot 500m isobath
  geom_contour(data = bathy_df, aes(x = lon, y = lat, z = depth), breaks = -500, size = 0.25, colour = "black", lty = 3) +
  # Plot coastlines
  geom_polygon(data = coastlines_10m, aes(x = lon, y = lat, group = group), fill = "grey70") +
  # Areas of interest
  geom_shape(data = area_BF_CAA_2, aes(x = x,y = y), fill = NA, col = "black", lwd = 0.2, lty = 1) +
  geom_shape(data = area_BB_2, aes(x = x,y = y), fill = NA, col = "black", lwd = 0.2, lty = 1) +
  geom_shape(data = area_SV_2, aes(x = x,y = y), fill = NA, col = "black", lwd = 0.2, lty = 1) +
  # Arctic circle
  geom_segment(aes(x = -180, xend = 100, y = 66.5, yend = 66.5), col = "#666766", lty = 2, size = 0.3) +
  # Plot data
  geom_tile(aes(fill = mean_ano_sa_int_d), color = "grey30", alpha = 0.8, size = 0.15) +
  scale_fill_manual("2015-2017\nAverage\nbackscatter\nanomaly", values = col_pal) +
  coord_map("azequalarea", ylim = c(66, 90), xlim = c(-160, 30), orientation = c(90, 0, -60)) +
  # Theme
  guides(fill = guide_legend(keywidth = 0.2, keyheight = 0.2, default.unit = "in"), color = "none") +
  theme(legend.position = "right", 
        legend.title = element_text(hjust = 0.5),
        panel.border = element_rect(fill = NA),
        strip.text.x = element_blank(), 
        strip.background = element_blank(), 
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank())
map_sa_anomaly
```

## Table 1. Abundance mesopelagic fish

```{r table1-heatmap-abundance, fig.align='center', fig.width=4.5, fig.height=3.25, dpi=600}
relative_abundance_area <- trawl_station %>%
  group_by(area) %>%
  summarise(total_fish = sum(number_fish)) %>% # Calculate total number of fish caught per area
  ungroup() %>%
  right_join(., trawl_area, by = c("area")) %>% # Join with relative abundance dataset
  mutate(area = factor(case_when(area == "BF_CAA" ~ "Beaufort Sea &\nCan. Arct. Archipelago",
                                 area == "BB" ~ "Baffin Bay",
                                 area == "SV" ~ "Svalbard"),
                       levels = c("Beaufort Sea &\nCan. Arct. Archipelago", "Baffin Bay", "Svalbard"))) %>%
  unite(area_label, area, total_fish, sep = "\nn = ", remove = F) %>% # Create labels for x axis
  # Format species name appropriately
  mutate(fish_species = str_replace(fish_species, pattern = "_", replacement = " "),
         fish_species = str_replace(fish_species, pattern = " sp", replacement = " sp."),
         fish_species = str_to_sentence(fish_species),
         fish_species = if_else(fish_species == "Notolepis rissoi", "Arctozenus risso", fish_species),
         fish_species = factor(fish_species, levels = c("Icelus bicornis",
                                                        "Anarhichas lupus",
                                                        "Reinhardtius hippoglossoides",
                                                        "Arctozenus risso",
                                                        "Melanogrammus aeglefinus",
                                                        "Gadus morhua",
                                                        "Leptoclinus maculatus",
                                                        "Sebastes sp.",
                                                        "Myctophidae",
                                                        "Liparidae",
                                                        "Boreogadus saida")),
         color_text = if_else(RA_area >= 40, "white", "black"),
         RA_area_round = 10 * floor(RA_area / 10),
         area_label = factor(area_label, levels = c("Beaufort Sea &\nCan. Arct. Archipelago\nn = 923", 
                                                    "Baffin Bay\nn = 232",
                                                    "Svalbard\nn = 464"))) %>%  
  # Plot
  ggplot() +
  geom_tile(aes(x = area_label, y = fish_species, fill = RA_area_round), color = "white", lwd = 0.6) +
  geom_text(aes(x = area_label, y = fish_species, label = sprintf("%0.1f", round(RA_area, 1)), color = color_text), size = 3) +
  scale_x_discrete(expand = c(0, 0)) +
  scale_y_discrete(labels = c("Twohorn sculpin",
                              "Atlantic wolffish",
                              "Greenland halibut",
                              "Spotted barracudina",
                              "Haddock", 
                              "Atlantic cod",
                              "Daubed shanny",
                              "Redfish",
                              "Lanternfish",
                              "Snailfish",
                              "Polar cod"), 
                   expand = c(0, 0)) +
  scale_colour_identity() +
  scale_fill_viridis_c("Relative\nabundance (%)  ", option = "mako", breaks = seq(0, 100, 20), limits = c(0, 100), direction = -1) +
  guides(fill = guide_colorbar(label.position = "top")) +
  theme(axis.title = element_blank(), 
        legend.position = "top", 
        legend.title = element_text(vjust = 0), 
        legend.title.align = 0.5, 
        legend.key.height = unit(0.1, "in"), 
        legend.key.width = unit(0.3, "in"),
        legend.box.margin = margin(0,0,-5,0))
relative_abundance_area
```


# Supplementary data

## Table S3. Acoustic settings

```{r tableS3-acoustic-settings, echo=F, warning=F, message=F, fig.align='center'}
summary_MVBS <- MVBS %>%
  group_by(date,process_ID, filename, area, year) %>%
  summarise() %>% # Remove depth component
  ungroup() %>%
  group_by(area,year) %>% # Summarise total amount of files and total number of hours of data
  summarise(n_files=n(),
            duration_min=n_files*10,
            duration_h=round(duration_min/60))

# Create data frame and table
data.frame(year=c(rep(2015,3), rep(2016,3), rep(2017,3)),
           Area=rep(c("Beaufort Sea","Baffin Bay","Svalbard"), 3),
           area=rep(c("BF_CAA","BB","SV"),3),
           Vessel=c("CCGV Amundsen","CCGV Amundsen","RV Polarstern", # 2015
                    "CCGV Amundsen","CCGV Amundsen","RV Helmer Hanssen", # 2016
                    "FV Frosti","CCGV Amundsen","RV Polarstern"), # 2017
           Echosounder=c(rep("EK60",6), rep("EK80",1), rep("EK60",2)),
           Transducer_depth=c(rep(7,2),11,rep(7,2),8,6,7,11),
           Pulse_length=rep(1.024,9),
           Power=c(rep(2000,2),1000,rep(2000,5),1000)) %>%
  left_join(.,summary_MVBS, by=c("area", "year")) %>%
  dplyr::select(year, Area, Vessel, Echosounder, Transducer_depth, Pulse_length,
                Power, duration_h) %>%
  rename("Year"=year,
         "Transducers depth (m)"=Transducer_depth, 
         "Pulse length (msec)"=Pulse_length,
         "Power (W)"=Power,
         "Acoustic data duration (h)"=duration_h) %>%
  mutate(Year="" ) %>% # remove year values for a more readable table
  kbl(align=c("l", "l", "l", "c", "c","c","c","c")) %>%
  kable_classic(full_width=F, html_font="Arial") %>%
  pack_rows("2015", 1, 3) %>%
  pack_rows("2016", 4, 6) %>%
  pack_rows("2017", 7, 9)
```

