---
title: "Pan-Arctic DSL - Arctic current velocities and directions"
author: "Pierre Priou - pierre.priou@mi.mun.ca"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(dtplyr)
library(lubridate)
library(plotly)
library(cowplot)
library(raster)
library(sf)
library(metR)
library(cmocean)
library(tidync)
```

```{r figure-theme}
theme_set(theme_bw())
theme_update(panel.border = element_blank(), 
             axis.line = element_line(),
             axis.text = element_text(size = 10),
             axis.title = element_text(size = 11),
             legend.title = element_text(size = 11),
             strip.text.x = element_text(size = 11, face = "plain", 
                                         hjust = 0.5),
             strip.background = element_rect(colour = "transparent", 
                                             fill = "transparent"),
             legend.margin = margin(0, 0, 0, 0),
             legend.box.margin = margin(0, 0, -8, 0),
             plot.margin = unit(c(0.075, 0.075, 0.075, 0.075), "in"))
```

# Aim

This document creates maps of annual mean current velocity and directions in 2015, 2016, and 2017. 

# NetCDF to csv

First I convert Arctic Ocean Physics Reanalysis data from NetCDF to csv files. 

```{r NetCDF-csv, eval=FALSE}
# Empty dataframe
velo <- data.frame() 

# List files per year
file_list <- list.files("data/ocean_currents/nc", pattern = "*.nc", full.names = T) 

# Loop through each file
for (i in file_list) { 
  
  # Temperature, salinity, depth, and current velocity
  velo_tmp <- tidync(i) %>%
    activate("D0,D1,D3,D2") %>% 
    hyper_tibble()
  
  # Coordinates
  coord_tmp <- tidync(i) %>%
    activate("D0,D1") %>% 
    hyper_tibble() %>%
    # Find extent of grid
    mutate(xmin = min(x),
           xmax = max(x),
           ymin = min(y), 
           ymax = max(y))
  
  # Retrieve projection details
  r <- terra::rast(i)
  proj_tmp <- tibble(proj = crs(r, proj = T),
                     res_x = r@ptr$res[1],
                     res_y = r@ptr$res[2])

  # Combine datasets
  tmp <- left_join(velo_tmp, coord_tmp, by = c("x", "y")) %>%
    # Add projection details
    bind_cols(., proj_tmp) %>%
    # Calculate date and add extents
    mutate(date = as.POSIXct(time * 3600, origin = "1950-01-01 00:00:00", tz = "UTC")) %>%
    # Rename variables
    rename(lat = latitude, 
           lon = longitude) %>%
    # Only keep velocity data 
    dplyr::select(date, lat, lon, x, y, depth, vxo, vyo, proj, res_x, res_y,
                  xmin, xmax, ymin, ymax)
  
  # Append data in a single file
  velo <- bind_rows(velo, tmp)
  print(paste0("Processing of ", i, " over."))
}

# Write csv file
write_csv(velo, file = "data/ocean_currents/velocity_318m_CMEMS.csv")

rm(velo, file_list, i, velo_tmp, coord_tmp, r, proj_tmp, tmp)
```

# Data tidying

I load the csv file and tidy it. 

```{r load-data}
# List files
velo_raw <- read_csv("data/ocean_currents/velocity_318m_CMEMS.csv", show_col_types = F)

# Get projection, cell resolution, and extents of data
proj_current <- velo_raw$proj[1]
res_current <- c(velo_raw$res_x[1], velo_raw$res_x[1])
extent_current <- c(velo_raw$xmin[1], velo_raw$xmax[1], velo_raw$ymin[1], velo_raw$ymax[1])

# Remove projection variables from tibble
velo <- velo_raw %>%
  dplyr::select(-proj, -res_x, -res_y) %>%
  # Add month and year, and calculate velocity (in cm/s)
  mutate(year = year(date),
         month = month(date),
         vxo = vxo * 100,
         vyo = vyo * 100,
         velocity = sqrt(vxo^2 + vyo^2))
```

The original projection of the data is: *`r proj_current`*

Each pixel represents a `r paste0(res_current[1] * 100, " km x ", res_current[2] * 100, " km")` cell.

The variables are:

-   time: time in hours since 1950-01-01 00:00:00
-   latitude: latitude
-   longitude: longitude
-   x: x-coordinate (polar stereographic x coordinate)
-   y: y-coordinate (polar stereographic y coordinate)
-   vxo: Zonal velocity m/s (x)
-   vyo: Meridional velocity m/s (y)

```{r load-coastline-data, message=FALSE, warning=FALSE, include=FALSE}
proj_coast <- raster(extent(extent_current), crs = proj_current, resolution = res_current)

# Coastline shapefile
coast <- rgdal::readOGR("data/bathy/ne_10m_land.shp", verbose = F) %>% 
  # spTransform(CRSobj = crs("EPSG:4326")) %>% 
  crop(extent(-180, 180, 0, 90)) %>%
  # Project shapefile in new proj
  spTransform(CRSobj = crs(proj_current)) %>% 
  # Convert to a dataframe for ggplot
  fortify() %>% 
  rename(x = long, y = lat) %>%
  mutate(x = x * 10^-5,
         y = y * 10^-5)
```

Here is a plot of current velocity at 318 m depth in October 2016.

```{r plot-velocity-oct-2016-12.5km, message=FALSE, fig.height=8.5, fig.width=10}
velo %>%
  group_by(year, x, y) %>%
  # Calculate annual mean
  summarise(vxo = mean(vxo),
            vyo = mean(vyo)) %>%
  ungroup() %>%
  mutate(velocity = sqrt(vxo^2 + vyo^2)) %>%
  # Only plot 2016
  filter(year == 2016) %>%
  ggplot(aes(x = x, y = y)) +
  geom_tile(aes(fill = velocity)) +
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  scale_fill_cmocean(name = "delta", limits = c(0, 10), oob = scales::squish, alpha = 0.5) +
  coord_fixed(xlim = c(-27, 27), ylim = c(-35, 18), expand = F) +
  guides(mag = "none", alpha = "none") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
```

# Reduce data resolution

Define projections. 

```{r current-projection}
# Set current projection (12.5 x 12.5 km)
r_current <- raster(extent(extent_current), crs = proj_current, resolution = res_current)

# Set new projection 
res_new <- c(150, 150) # in km
r_new <- raster(extent(extent_current), crs = proj_current, resolution = res_new / 100)
```

Calculate the mean velocity and direction for each grid cell between 2015 and 2017.

```{r rasterize-data}
# Convert to spatial data
velo_new <- SpatialPointsDataFrame(
  SpatialPoints(cbind(velo$x, velo$y), proj4string = crs(r_current)),
  data.frame(lat = velo$lat,
            lon = velo$lon,
            vxo = velo$vxo,
            vyo = velo$vyo)) %>%
  # Rasterize on the new grid
  rasterize(., r_new, fun = mean, na.rm = T) %>% 
  # Remove ID layer
  dropLayer(1) %>% 
  # Convert raster to data frame
  rasterToPoints() %>% 
  as.data.frame() %>% 
  # Add year
  mutate(velocity = sqrt(vxo^2 + vyo^2))
```

The following plot show annual mean current velocity in 2016 with a `r paste0(res_new[1], " km x ", res_new[2], " km")` cell resolution.

```{r plot-velocity-oct-2016-new-res, message=FALSE, fig.height=3, fig.width=6.5}
map_velo <- velo_new %>%
  ggplot() +
  # geom_tile(aes(x = x, y = y, fill = velocity), col = NA) +
  geom_contour_fill(aes(x = x, y = y, z = velocity), binwidth = 0.5, col = NA) +
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  geom_vector(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity),
              size = 0.2, arrow.angle = 30, arrow.type = "open", arrow.length = 0.75) +
  scale_alpha(range = c(0.5, 0.8)) +
  scale_fill_cmocean(expression("Velocity\n(m s"^-1*")"), 
                     name = "delta", limits = c(0, 5), end = 0.7, oob = scales::squish, alpha = 0.5,
                     breaks = seq(0, 10, 1)) +
  coord_fixed(xlim = c(-27, 27), ylim = c(-40, 18), expand = F) +
  guides(mag = "none", alpha = "none") +
  theme(legend.position = "right",
        legend.key.height = unit(0.3, "in"),
        legend.key.width = unit(0.1, "in"),
        legend.title = element_text(hjust = 0.5),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_velo.png", map_velo, height = 3, width = 4, units = "in", dpi = 600)
```
