---
title: "PanArctic DSL - mid-water trawl data processing"
author: "Pierre Priou - pierre.priou@mi.mun.ca"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
---

```{r r-setup, message=FALSE, warning=FALSE} 
# Load packages
library(tidyverse) # Tidy code 
library(lubridate) # Date handling 
library(marmap) # Retrieve bottom depth 

# Suppress summarise() warning
options(dplyr.summarise.inform = F)
# Theme for figures
theme_set(theme_bw())
theme_update(panel.grid = element_blank(), 
             panel.border = element_blank(),
             axis.line = element_line(), 
             axis.text = element_text(size = 10), 
             axis.title = element_text(size = 11),
             legend.title = element_text(size = 11),
             strip.text.x = element_text(size = 11, face = "plain", hjust = 0.5),
             strip.background = element_rect(colour = "transparent", fill = "transparent"),
             legend.margin = margin(0, 0, 0, 0),
             legend.box.margin = margin(0, 0, -8, 0),
             plot.margin = unit(c(0.05, 0.05, 0.05, 0.05), "in"))
```

# Aim 

This document load and tidy mid-water trawl data from the RV *Helmer Hanssen*, FV *Frosti*, and CCGS *Amundsen*. Because the data is formatted differently between each vessel, I process the data per vessel and then combine the tidy data together. 

# Helmer Hanssen

We deployed a Harstad mid-water trawl with mesh size in the mesopelagic layer from the *Helmer Hanssen* in August 2016 (see Geoffroy et al. 2019 for more information). 

## Data loading

```{r HH-load-data}
# Load fish length
HH_length_raw <- read_csv("data/nets/Helmer_Hanssen_2016_fish_length.csv", show_col_types = F) %>%
  # Standardize species name
  mutate(fish_species = str_to_lower(fish_species)) %>%
  separate(fish_species, c("genus", "species"), remove = T) %>%
  unite(fish_species, genus, species, sep = "_", remove = F) %>%
  # Fix date
  unite(date, day, month, year, sep = "/", remove = F) %>%
  mutate(date = dmy(date),
         vessel = "Helmer Hanssen",
         deployment = "NA",
         tow_type = "MWT", 
         fish_ID = as.character(fish_numberID)) %>%
  rename(SL_mm = SL_mm_lab)

# Load abundance
HH_abund_raw <- read_csv("data/nets/Helmer_Hanssen_2016_fish_macrozooplankton.csv", show_col_types = F) %>%
  # Standardize species name
  mutate(fish_species = str_to_lower(fish_species)) %>%
  separate(fish_species, c("genus", "species"), remove = T) %>%
  unite(fish_species, genus, species, sep="_", remove = F) %>%
  # Fix date
  unite(date, day, month, year, sep="/", remove = F) %>%
  mutate(date = dmy(date),
         vessel = "Helmer Hanssen",
         deployment = "NA",
         number_fish = as.numeric(number_fish))
```

## Data tidying 

```{r HH-tidy-data} 
# Tidy length
HH_length <- HH_length_raw %>%
  dplyr::select(date, year, month, vessel, station, deployment, lat, lon, bottom_depth_trawl,
                max_sampling_depth, tow_type, fish_ID, fish_species, genus, species, SL_mm)

# Tidy abundance
HH_abund <- HH_abund_raw %>% 
  # Select fish 
  filter(nekton == T) %>% # Select relevant variables
  dplyr::select(date, year, month, vessel, station, deployment, lat, lon, bottom_depth_trawl, 
                max_sampling_depth, tow_type, fish_species, genus, species, number_fish, W_g)
```

# Amundsen

Several trawls and nets were deployed to groundtruth the acoustic signal. On the Amundsen a modified Isaac-Kidd Midwater Trawl (IKMT) was deployed from the A-frame at the bow of the ship in 2015 and 2016. Samples were either preserved indivdually or in bulk. For the total abundance individual samples need to be summed to the bulk samples. Length measurements were only taken on individual samples.

## Data loading

```{r AM-load-data}
# Load length of single fish
AM_length_raw <- read_csv("data/nets/Amundsen_2014_2019_nets_individuals_length.csv", show_col_types = F) %>%
  # Fix date
  unite(date, Sampling_date, Sampling_time, sep = " ", remove = F) %>%
  mutate(date = ymd_hm(format(mdy_hms(date, tz = "UTC"), format = '%Y%m%d %H:%M')),
         year = year(date),
         month = month(date),
         # Standardize species name
         family = str_to_lower(family),
         genus = str_to_lower(genus),
         species = str_to_lower(species),
         species = if_else(species == "sp.", "sp", species),
         genus = if_else(family == "liparidae" & genus == "liparis" & species == "sp", "liparidae", genus),
         genus = if_else(family == "liparidae" & genus == "-" & species == "sp", "liparidae", genus),
         genus = if_else(family == "liparidae" & is.na(genus) == T, "liparidae", genus),
         genus = if_else(family == "myctophidae" & genus == "-" & species == "sp", "myctophid", genus),
         genus = if_else(family == "myctophidae" & is.na(genus) == T & species == "sp", "myctophid", genus),
         genus = if_else(family == "myctophidae" & genus == "benthosema" & species == "glaciale", "myctophid", genus),
         species = if_else(family == "liparidae" & genus == "liparidae" & species == "sp", "uid", species),
         species = if_else(family == "liparidae" & genus == "liparidae" & is.na(species) == T, "uid", species),
         species = if_else(family == "myctophidae" & genus == "myctophid" & species == "sp", "uid", species),
         species = if_else(family == "myctophidae" & genus == "myctophid" & is.na(species) == T, "uid", species),
         species = if_else(family == "myctophidae" & genus == "myctophid" & species == "glaciale", "uid", species),
         gear = as.factor(str_to_lower(gear)),
         tow_type = as.factor(str_to_upper(tow_type)),
         number_fish_ind = 1,
         deployment = "NA",
         fish_ID = "NA",
         vessel = "Amundsen") %>%
  unite(fish_species, genus, species, sep = "_", remove = F) %>%
  unite(fish_ID, Year_Fish_id, Original_ID, sep="_", remove = F) %>%
  mutate() %>%
  # 1 date failed to parse because there is one empty row in the data (row 3709) so I removed it
  filter(is.na(date) == F) %>% 
  rename(SL_mm = SL_mm_lab)

# Load bulk samples
AM_bulk_raw <- read_csv("data/nets/Amundsen_2014_2019_nets_bulk_station.csv", show_col_types = F) %>%
  # Fix date
  unite(date, Sampling_date, Sampling_time, sep = " ", remove = F) %>%
  mutate(date = ymd_hm(format(mdy_hms(date, tz = "UTC"), format  ='%Y%m%d %H:%M')),
         year = year(date),
         month = month(date),
         station = as.character(station),
         # Standardize species name
         family = str_to_lower(family),
         genus = str_to_lower(genus),
         species = str_to_lower(species),
         species = if_else(species == "sp.", "sp", species),
         gear = as.factor(str_to_lower(gear)),
         tow_type = as.factor(str_to_upper(tow_type)),
         vessel = "Amundsen") %>%
  unite(fish_species, genus, species, sep = "_", remove = F) %>%
  # Rename number_fish to allow for merging with individual data
  rename(number_fish_bulk = number_fish) %>% 
  # Select relevant variables
  dplyr::select(date, year, month, vessel, station, lat, lon, bottom_depth_trawl,
                max_sampling_depth, tow_type, fish_species, genus, species, number_fish_bulk)
```

## Data tidying 

```{r AM-tidy-data} 
# Tidy length 
AM_length <- AM_length_raw %>% 
  filter(tow_type %in% c("IKMT", "IKRMT") & between(year, 2015, 2017)) %>%
  dplyr::select(date, year, month, vessel, station, deployment, lat, lon, bottom_depth_trawl,
                max_sampling_depth, tow_type, fish_ID, fish_species, genus, species, SL_mm)

# Tidy abundance
AM_abund <- AM_length_raw %>% # Summarise data per station 
  group_by(date, year, month, vessel, station, deployment, lat, lon, bottom_depth_trawl,
           max_sampling_depth, tow_type, fish_species, genus, species) %>%
  summarise(number_fish_ind = sum(number_fish_ind)) %>%
  ungroup() %>% 
  # Add bulk data
  left_join(., AM_bulk_raw,
            by = c("date", "year", "month", "vessel", "station", "lat", "lon", "bottom_depth_trawl",
                    "max_sampling_depth", "tow_type", "fish_species", "genus", "species")) %>%
  # Sum individual and bulk fish 
  rowwise() %>% 
  mutate(number_fish = sum(number_fish_ind, number_fish_bulk, na.rm = T)) %>% 
  # Select IKMT data
  filter(tow_type %in% c("IKMT", "IKRMT") & between(year, 2015, 2017)) %>% 
  # Remove unused variables
  dplyr::select(-number_fish_ind, -number_fish_bulk)
```

# Frosti

On the FV _Frosti_ a mid-water otter trawl was deployed in 2107 in the Beaufort Sea. 

```{r load-bathy-BF}
# Bathy data
BF_bathy <- getNOAA.bathy(lon1 = -142, lon2 = -115, lat1 = 68, lat2 = 73, resolution = 1, 
                          keep = T, path = "data/bathy/")
```

## Data loading

The Frosti did not recorded bottom depth of the deployment, so I find it using ETOPO bathymetry.

```{r FI-data-loading} 
# Load metadata
FI_metadata <- read_csv("data/nets/Frosti_2017_net_metadata.csv", show_col_types = F) %>%
  # Fix variables 
  mutate(date = mdy(Date_UTC),
         year = year(date), 
         month = month(date),
         Fish_Sheet_Station_ID = str_to_lower(Fish_Sheet_Station_ID),
         station = str_replace(Fish_Sheet_Station_ID, pattern = "-", replacement = ""), 
         Fish_Sheet_Deployment_ID = str_to_lower(Fish_Sheet_Deployment_ID),
         deployment = str_replace(Fish_Sheet_Deployment_ID, pattern = "-", replacement = ""), 
         tow_type = str_to_lower(Gear), 
         # Match names between metadata and data 
         tow_type = if_else(tow_type == "mid-water", "mwt", tow_type),
         tow_type = factor(str_to_upper(tow_type)),
         max_sampling_depth = as.numeric(Target_depth_m)) %>% 
  arrange(date) %>%
  # Calculate mid coordinates for the deployment
  rowwise() %>%
  mutate(lat = if_else(is.na(End_Latitude) == T, Start_Latitude, mean(c(Start_Latitude, End_Latitude), na.rm = T)),
         lon = if_else(is.na(End_Longitude) == T, Start_Longitude, mean(c(Start_Longitude, End_Longitude), na.rm = T)), 
         # Find bottom depth using bathymetry data 
         get_depth = get.depth(BF_bathy, x = lon, y = lat, locator = F), 
         bottom_depth_trawl = as.numeric(abs(get_depth$depth))) %>% 
  # Select relevant variables 
  dplyr::select(date, year, month, station, deployment, lat, lon, tow_type, max_sampling_depth, bottom_depth_trawl)

FI_length_raw <- read_csv("data/nets/Frosti_2017_fish_weights_lengths.csv", show_col_types = F) %>%
  mutate(year = Year, 
         tow_type = str_to_lower(Gear),
         # Match names between metadata and data
         tow_type = if_else(tow_type == "wa2", "w2a", tow_type),
         tow_type = factor(str_to_upper(tow_type)),
         station = str_to_lower(Station),
         station = str_replace(station, pattern = "-", replacement = ""),
         deployment = str_to_lower(Deployment),
         deployment = str_replace(deployment, pattern = "-", replacement = ""),
         fish_ID = as.character(Fish_ID), family = str_to_lower(Family),
         genus = str_to_lower(Genus),
         genus = if_else(family == "liparidae" & is.na(genus) == T, "liparidae", genus),
         species = str_to_lower(Species), 
         species = if_else(family == "liparidae" & genus == "liparidae" & is.na(species) == T, "uid", species),
         number_fish = if_else(is.na(Number_in_bag) == T, 1, Number_in_bag),
         W_g = Weight_g, 
         SL_mm = as.numeric(Standard_Length_mm_lab),
         vessel = "Frosti") %>% 
  unite(fish_species, genus, species, sep = "_", remove = F) %>% 
  arrange(year)
```

## Tidy data

```{r FI-tidy-data}
# Tidy length
FI_length <- FI_length_raw %>%
  # Add metadata
  left_join(., FI_metadata, by = c("year", "tow_type", "station", "deployment")) %>%
  # Select data
  filter(tow_type == "MWT" & year == 2017) %>%
  # Select variables of interest
  dplyr::select(date, year, month, vessel, station, deployment, lat, lon, bottom_depth_trawl, 
                max_sampling_depth, tow_type, fish_ID, fish_species, genus, species, SL_mm) 

# Tidy abundance
FI_abund <- FI_length_raw %>%
  # Add metadata
  left_join(., FI_metadata, by = c("year", "tow_type", "station", "deployment")) %>%
  # Select data
  filter(tow_type == "MWT" & year == 2017) %>%
  # Calculate abundance per station
  group_by(date, year, month, vessel, station, deployment, lat, lon, bottom_depth_trawl, 
                max_sampling_depth, tow_type, fish_species, genus, species) %>%
  summarise(number_fish = sum(number_fish)) %>%
  ungroup()
```

# Combine data

Now I combine data from all ships in a single file. I select tows that were performed below 200 m sampling depth.

```{r combine-trawls} 
# Length 
length_station <- bind_rows(HH_length, AM_length, FI_length) %>%
  filter(max_sampling_depth >= 200)

# Abundance
abund_station <- bind_rows(HH_abund, AM_abund, FI_abund) %>% 
  filter(max_sampling_depth >= 200)
```

I save the data as `.csv` and `.RData` files.

```{r save-data}
# Save .csv
write_csv(length_station, file = "data/nets/fish_length_station.csv")
write_csv(abund_station, file = "data/nets/fish_abundance_station.csv")

# Save .Rdata
save(length_station, file = "data/nets/fish_length_station.RData")
save(abund_station, file = "data/nets/fish_abundance_station.RData")
```