---
title: "Pan-Arctic DSL - Arctic current velocities and directions"
author: "Pierre Priou - pierre.priou@mi.mun.ca"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: show
    toc: true
    toc_float: true
    toc_collapsed: true
---

```{r load-packages, message=FALSE}
library(tidyverse)
library(dtplyr)
library(lubridate)
library(plotly)
library(cowplot)
library(raster)
library(sf)
library(metR)
library(cmocean)
library(tidync)
library(marmap)
source("R/getNOAA.ice.bathy.R") # Load bathy data from NOAA
```

```{r figure-theme}
theme_set(theme_bw())
theme_update(panel.border = element_blank(), 
             axis.line = element_line(),
             axis.text = element_text(size = 10),
             axis.title = element_text(size = 11),
             legend.title = element_text(size = 11),
             strip.text.x = element_text(size = 11, face = "plain", 
                                         hjust = 0.5),
             strip.background = element_rect(colour = "transparent", 
                                             fill = "transparent"),
             legend.margin = margin(0, 0, 0, 0),
             legend.box.margin = margin(0, 0, -8, 0),
             plot.margin = unit(c(0.075, 0.075, 0.075, 0.075), "in"))
```

# Aim

This document creates maps of annual mean current velocity and directions, and temperature in 2015, 2016, and 2017. 

# NetCDF to csv

First I convert Arctic Ocean Physics Reanalysis data from NetCDF to csv files. The raw file contains these variables:

-   time: time in hours since 1950-01-01 00:00:00
-   latitude: latitude
-   longitude: longitude
-   x: x-coordinate (polar stereographic x coordinate)
-   y: y-coordinate (polar stereographic y coordinate)
-   vxo: Zonal velocity m/s (x)
-   vyo: Meridional velocity m/s (y)

The csv file also contains the date.

```{r NetCDF-csv, eval=FALSE}
# Empty dataframe
velo <- data.frame() 

# List files per year
file_list <- list.files("data/ocean_currents/nc", pattern = "380", full.names = T) 

# Loop through each file
for (i in file_list) { 
  
  # Temperature, salinity, depth, and current velocity
  velo_tmp <- tidync(i) %>%
    activate("D0,D1,D3,D2") %>% 
    hyper_tibble()
  
  # Coordinates
  coord_tmp <- tidync(i) %>%
    activate("D0,D1") %>% 
    hyper_tibble() %>%
    # Find extent of grid
    mutate(xmin = min(x),
           xmax = max(x),
           ymin = min(y), 
           ymax = max(y))
  
  # Retrieve projection details
  r <- terra::rast(i)
  proj_tmp <- tibble(proj = crs(r, proj = T),
                     res_x = r@ptr$res[1],
                     res_y = r@ptr$res[2])

  # Combine datasets
  tmp <- left_join(velo_tmp, coord_tmp, by = c("x", "y")) %>%
    # Add projection details
    bind_cols(., proj_tmp) %>%
    # Calculate date and add extents
    mutate(date = as.POSIXct(time * 3600, origin = "1950-01-01 00:00:00", tz = "UTC")) %>%
    # Rename variables
    rename(lat = latitude, 
           lon = longitude) %>%
    # Only keep velocity data 
    dplyr::select(date, lat, lon, x, y, depth, vxo, vyo, thetao, proj, res_x, res_y,
                  xmin, xmax, ymin, ymax)
  
  # Append data in a single file
  velo <- bind_rows(velo, tmp)
  print(paste0("Processing of ", i, " over."))
}

# Write csv file
write_csv(velo, file = "data/ocean_currents/velocity_380m_CMEMS.csv")

rm(velo, file_list, i, velo_tmp, coord_tmp, r, proj_tmp, tmp)
```

# Data tidying

I load the csv file and calculate velocity and angle. 

```{r load-data}
# List files
velo_raw <- read_csv("data/ocean_currents/velocity_380m_CMEMS.csv", show_col_types = F)

# Get projection, cell resolution, and extents of data
proj_current <- velo_raw$proj[1]
res_current <- c(velo_raw$res_x[1], velo_raw$res_x[1])
extent_current <- c(velo_raw$xmin[1], velo_raw$xmax[1], velo_raw$ymin[1], velo_raw$ymax[1])

# Remove projection variables from tibble
velo <- velo_raw %>%
  dplyr::select(-proj, -res_x, -res_y) %>%
  # Add month and year, and calculate velocity (in cm/s) and angle (in radians)
  mutate(year = year(date),
         month = month(date),
         vxo = vxo * 100,
         vyo = vyo * 100,
         velocity = sqrt(vxo^2 + vyo^2),
         angle_deg = ((atan((vyo / vxo)) * 180) / pi) - 180)
```

The original projection of the data is: *`r proj_current`*

Each pixel represents a `r paste0(res_current[1] * 100, " km x ", res_current[2] * 100, " km")` cell.

```{r load-coastline-data, message=FALSE, warning=FALSE, include=FALSE}
proj_coast <- raster(extent(extent_current), crs = proj_current, resolution = res_current)

# Coastline shapefile
coast <- rgdal::readOGR("data/bathy/ne_10m_land.shp", verbose = F) %>% 
  # spTransform(CRSobj = crs("EPSG:4326")) %>% 
  crop(extent(-180, 180, 0, 90)) %>%
  # Project shapefile in new proj
  spTransform(CRSobj = crs(proj_current)) %>% 
  # Convert to a dataframe for ggplot
  fortify() %>% 
  rename(x = long, y = lat) %>%
  mutate(x = x * 10^-5,
         y = y * 10^-5)

# Arctic circle and areas of interest
arctic_circle <- data.frame(lon = seq(-180, 180, 0.1)) %>%
  mutate(lat = 66.5) %>%
  st_as_sf(coords = c("lon", "lat"), remove = F, crs = st_crs(4326)) %>%
  st_transform(crs = st_crs(proj_coast)) %>%
  mutate(x = st_coordinates(.)[,1] * 10^-5,
         y = st_coordinates(.)[,2] * 10^-5) %>%
  st_drop_geometry()
```

Here is a plot of current velocity at 380 m depth averaged over 2015-2017.

```{r plot-velocity-12.5km, message=FALSE, fig.height=8.5, fig.width=10}
velo %>%
  group_by(x, y) %>%
  # Calculate annual mean
  summarise(vxo = mean(vxo),
            vyo = mean(vyo)) %>%
  ungroup() %>%
  mutate(velocity = sqrt(vxo^2 + vyo^2)) %>%
  ggplot(aes(x = x, y = y)) +
  geom_tile(aes(fill = velocity, col = velocity)) +
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  scale_color_cmocean(name = "speed", limits = c(0, 10), oob = scales::squish,) +
  scale_fill_cmocean(name = "speed", limits = c(0, 10), oob = scales::squish,) +
  coord_fixed(xlim = c(-27, 27), ylim = c(-35, 18), expand = F) +
  guides(mag = "none", alpha = "none") +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
```

Plot of temperature at 380 m depth.

```{r plot-temperature-12.5km, message=FALSE, fig.height=8.5, fig.width=10}
map_temp <- velo %>%
  group_by(x, y) %>%
  # Calculate annual mean
  summarise(thetao = mean(thetao)) %>%
  ungroup() %>%
  ggplot(aes(x = x, y = y)) +
  geom_tile(aes(fill = thetao, col = thetao)) +
  geom_contour(aes(z = thetao), breaks = 0, col = "green", lwd = 0.25) +
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  scale_color_cmocean("T", name = "thermal", limits = c(-1, 5), oob = scales::squish) +
  scale_fill_cmocean("T", name = "thermal", limits = c(-1, 5), oob = scales::squish) +
  coord_fixed(xlim = c(-40, 40), ylim = c(-40, 40), expand = F) +
  theme(legend.position = "right",
        legend.key.height = unit(0.3, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_temp.png", map_temp, height = 5, width = 5, units = "in", dpi = 600)
map_temp
```

# Reduce data resolution

Define projections and calculate the mean velocity and direction for each grid cell between 2015 and 2017.

```{r current-projection-raster}
# Set current projection (12.5 x 12.5 km)
r_current <- raster(extent(extent_current), crs = proj_current, resolution = res_current)

# Set new projection 
res_new <- c(125, 125) # in km
r_new <- raster(extent(extent_current), crs = proj_current, resolution = res_new / 100)

# Convert to spatial data
velo_new <- SpatialPointsDataFrame(
  SpatialPoints(cbind(velo$x, velo$y), proj4string = crs(r_current)),
  data.frame(lat = velo$lat,
             lon = velo$lon,
             vxo = velo$vxo,
             vyo = velo$vyo,
             thetao = velo$thetao)) %>%
  # Rasterize on the new grid
  rasterize(., r_new, fun = mean, na.rm = T) %>% 
  # Remove ID layer
  dropLayer(1) %>% 
  # Convert raster to data frame
  rasterToPoints() %>% 
  as.data.frame() %>% 
  # Add year
  mutate(velocity = sqrt(vxo^2 + vyo^2)) %>%
  complete(x = seq(min(x), max(x), res_new[1] / 100), y = seq(min(y), max(y), res_new[2] / 100))
```

Do the same thing at much finer grid resolution. This is used to plot the actual bathymetric contour on the map and not the large cells.

```{r new-projection-raster-HD}
# Set new projection 
res_new_HD <- c(25, 25) # in km
r_new_HD <- raster(extent(extent_current), crs = proj_current, resolution = res_new_HD / 100)

# Convert to spatial data
velo_new_HD <- SpatialPointsDataFrame(
  SpatialPoints(cbind(velo$x, velo$y), proj4string = crs(r_current)),
  data.frame(lat = velo$lat,
             lon = velo$lon,
             vxo = velo$vxo,
             vyo = velo$vyo,
             thetao = velo$thetao)) %>%
  # Rasterize on the new grid
  rasterize(., r_new_HD, fun = mean, na.rm = T) %>% 
  # Remove ID layer
  dropLayer(1) %>% 
  # Convert raster to data frame
  rasterToPoints() %>% 
  as.data.frame() %>% 
  # Add year
  mutate(velocity = sqrt(vxo^2 + vyo^2)) %>%
  complete(x = seq(min(x - 2 * (res_new_HD[1] / 100)), max(x + 2 * (res_new_HD[2] / 100)), res_new_HD[1] / 100), 
           y = seq(min(y - 2 * (res_new_HD[2] / 100)), max(y + 2 * (res_new_HD[2] / 100)), res_new_HD[2] / 100))
```

I also superimpose the grid cells of the GAM on the oceanic circulation map. First I load GAM data and re project their coordinates in the current velocity projection.

```{r load-and-project-GAM}
load("data/statistics/GAMSA_results.RData")
rm(GAMSA_I3, GAMSA_S3, predSA_I, predSA_S)

# Laea projection for bathy and all other files
arctic_laea <- raster(extent(-3500, 3500, -3500, 3500), crs = "EPSG:6931") 
projection(arctic_laea) <- gsub("units=m", "units=km", # Convert from m to km
                                projection(arctic_laea)) 
res(arctic_laea) <- c(150, 150) # Define cell resolution

# Project GAM cells onto the current projection
SA_new <- SA_df %>%
  group_by(xc, yc) %>%
  summarise() %>%
  ungroup() %>%
  # Convert to spatial data
  st_as_sf(coords = c("xc", "yc"), remove = F, crs = st_crs(6931)) %>%
  # Change projection
  st_transform(crs = st_crs(r_new)) %>%
  # Extract coordinates
  # /100 to match currents. Probably due to changes in units not accounted for during re-projection
  mutate(x = st_coordinates(.)[,1] / 100,
         y = st_coordinates(.)[,2] / 100) %>%
  st_drop_geometry()
```

# Plots

## Arctic map

The following plot show annual mean current velocity in 2016 with a `r paste0(res_new[1], " km x ", res_new[2], " km")` cell resolution.

```{r plot-velocity-new-res, warning=FALSE, fig.height=3, fig.width=4}
map_velo <- velo_new %>%
  ggplot() +
  # Plot interpolated current velocity
  geom_contour_fill(aes(x = x, y = y, z = velocity), binwidth = 0.5, na.fill = T) +
  # Cover areas that are shallower than our data
  geom_tile(data = subset(velo_new_HD, is.na(vxo) == T), aes(x = x, y = y), fill = "white", col = "white") +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  annotate("rect", xmin = 3, xmax = 15, ymin = -10, ymax = 3, fill = NA, col = "#F7941D", lwd = 0.5) +
  annotate("rect", xmin = -12, xmax = 0, ymin = -27, ymax = -7.5, fill = NA, col = "#F7941D", lwd = 0.5) +
  annotate("rect", xmin = -24, xmax = -13, ymin = -10, ymax = 6, fill = NA, col = "#F7941D", lwd = 0.5) +
  # Current vectors
  geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), size = 0.2, arrow.angle = 30,
             arrow.type = "open", arrow.length = 0.5) +
  # Arctic circle
  geom_path(data = arctic_circle, aes(x = x, y = y), col = "white", lty = 2, size = 0.3) +
  # Plot North Pole
  geom_point(aes(x = 0, y = 0), shape = 4, size = 0.7, stroke = 0.3, col = "white") +
  # GAM locations
  # geom_tile(data = SA_new, aes(x = x, y = y), fill = NA, colour = "red", size = 3.8) + 
  # Scales
  scale_fill_viridis_c("V", option = "viridis", limits = c(0, 5), end = 1, oob = scales::squish,
                       alpha = 0.5, breaks = seq(0, 10, 0.5)) +
  scale_x_continuous(breaks = seq(-50, 50 , 5)) +
  scale_y_continuous(breaks = seq(-50, 50 , 5)) +
  coord_fixed(xlim = c(-28, 28), ylim = c(-34, 20), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "right",
        legend.key.height = unit(0.3, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_velo.png", map_velo, height = 8, width = 8, units = "in", dpi = 600)
# map_velo
```

I print several layers of the plot for easy implementation in Illustrator.

Base map.

```{r base-map, warning=FALSE, fig.height=3, fig.width=4}
map_base <- velo_new_HD %>%
  ggplot() +
  # Plot interpolated current velocity
  geom_contour_fill(aes(x = x, y = y, z = velocity), binwidth = 0.5, na.fill = T) +
  # Cover areas that are shallower than our data
  geom_tile(data = subset(velo_new_HD, is.na(vxo) == T), aes(x = x, y = y), fill = "white", col = "white") +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  # Current vectors
  # geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), size = 0.2, arrow.angle = 30,
  #            arrow.type = "open", arrow.length = 0.5) +
  # Arctic circle
  geom_path(data = arctic_circle, aes(x = x, y = y), col = "white", lty = 2, size = 0.3) +
  # Plot North Pole
  # geom_point(aes(x = 0, y = 0), shape = 3, size = 0.7, stroke = 0.7, col = "#F05223") +
  # GAM locations
  # geom_tile(data = SA_new, aes(x = x, y = y), fill = NA, colour = "red", size = 3.6) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.8, 
                       oob = scales::squish, alpha = 0.7, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(-45, 45), ylim = c(-45, 45), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "right",
        legend.key.height = unit(0.3, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_velo_base.png", map_base, height = 5, width = 5, units = "in", dpi = 600)
```

Current vectors only. 

```{r map-current-vectors, warning=FALSE, fig.height=3, fig.width=4}
map_vectors <- velo_new %>%
  # filter(between(x, -40, 40) & between(y, -45, 40)) %>%
  ggplot() +
  # Plot interpolated current velocity
  geom_contour_fill(aes(x = x, y = y, z = velocity), binwidth = 0.5, na.fill = T) +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "transparent") +
  # Current vectors
  geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), skip = 0, size = 0.15, 
             arow.angle = 30, arrow.type = "open", arrow.length = 0.75) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.7, oob = scales::squish,
                     alpha = 0, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(-45, 45), ylim = c(-45, 45), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "right",
        legend.key.height = unit(0.3, "in"),
        legend.key.width = unit(0.1, "in"),
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent", colour = NA),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_vectors.png", map_vectors, bg = "transparent", height = 5, width = 5, units = "in", dpi = 600)
```

## Regional maps

For regional maps, I remove data that is outside of the areas of interest. 

```{r projection-HD-regions}
# Convert to spatial data
velo_HD <- velo_new_HD %>%
  # Remove data that is not relevant for regional maps
  mutate(vxo = if_else(between(x, -Inf, Inf) & between(y, -Inf, -31), NaN,
               if_else(between(x, 0, Inf) & between(y, -Inf, -14), NaN, vxo)),
               # if_else(between(x, 0, Inf) & between(y, 4, Inf), NaN,
               # if_else(between(x, -12, 2) & between(y, -6, Inf), NaN, vxo)))), 
         vyo = if_else(between(x, -Inf, Inf) & between(y, -Inf, -31), NaN,
               if_else(between(x, 0, Inf) & between(y, -Inf, -14), NaN, vyo)))
               # if_else(between(x, 0, Inf) & between(y, 4, Inf), NaN, 
               # if_else(between(x, -12, 2) & between(y, -6, Inf), NaN, vyo)))))
  
```

Northern Barents Sea.

```{r map-BA, warning=FALSE}
# Map limits
xmin <- -1
xmax <- 18
ymin <- -13
ymax <- 5

map_BA <- velo_HD %>%
 ggplot() +
  # Plot interpolated current velocity
  geom_contour_fill(aes(x = x, y = y, z = velocity), binwidth = 0.5, na.fill = T) +
  # Cover areas that are shallower than our data
  geom_tile(data = subset(velo_HD, is.na(vxo) == T), aes(x = x, y = y), fill = "white", col = "white") +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  # # Current vectors
  # geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), skip = 1, size = 0.15, arrow.angle = 30,
  #            arrow.type = "open", arrow.length = 0.5) +
  # Arctic circle
  geom_path(data = arctic_circle, aes(x = x, y = y), col = "white", lty = 2, size = 0.3) +
  # Plot North Pole
  geom_point(aes(x = 0, y = 0), shape = 4, size = 0.7, stroke = 0.3, col = "white") +
  # GAM locations
  # geom_tile(data = SA_new, aes(x = x, y = y), fill = NA, colour = "red", size = 9) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.8, 
                       oob = scales::squish, alpha = 0.7, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(xmin, xmax), ylim = c(ymin, ymax), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.line = element_blank(),    
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_velo_BA.png", map_BA, height = 2.5, width = 2.5, units = "in", dpi = 600)

map_BA_vector <- velo_HD %>%
 ggplot() +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "transparent") +
  # Current vectors
  geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), skip = 1, size = 0.2, arrow.angle = 30,
             arrow.type = "open", arrow.length = 0.5) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.8, 
                       oob = scales::squish, alpha = 0, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(xmin, xmax), ylim = c(ymin, ymax), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent", colour = NA),
        axis.line = element_blank(),    
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3, colour = NA))
ggsave("plots/map_velo_BA_vector.png", map_BA_vector, bg = "transparent",
       height = 2.5, width = 2.5, units = "in", dpi = 600)
```

Baffin Bay.

```{r map-BB, warning=FALSE}
# Map limits
xmin <- -15
xmax <- -1
ymin <- -30.5
ymax <- -10

map_BB <- velo_HD %>%
  ggplot() +
  # Plot interpolated current velocity
  geom_contour_fill(aes(x = x, y = y, z = velocity), binwidth = 0.5, na.fill = T) +
  # Cover areas that are shallower than our data
  geom_tile(data = subset(velo_HD, is.na(vxo) == T), aes(x = x, y = y), fill = "white", col = "white") +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  # Current vectors
  # geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), skip = 1, size = 0.15, arrow.angle = 30,
  #            arrow.type = "open", arrow.length = 0.5) +
  # Arctic circle
  geom_path(data = arctic_circle, aes(x = x, y = y), col = "white", lty = 2, size = 0.3) +
  # Plot North Pole
  geom_point(aes(x = 0, y = 0), shape = 4, size = 0.7, stroke = 0.3, col = "white") +
  # GAM locations
  # geom_tile(data = SA_new, aes(x = x, y = y), fill = NA, colour = "red", size = 12) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.8, 
                       oob = scales::squish, alpha = 0.7, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(xmin, xmax), ylim = c(ymin, ymax), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_velo_BB.png", map_BB, height = 3.55, width = 3.4, units = "in", dpi = 600)

map_BB_vector <- velo_HD %>%
 ggplot() +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "transparent") +
  # Current vectors
  geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), skip = 1, size = 0.2, arrow.angle = 30,
             arrow.type = "open", arrow.length = 0.5) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.8, 
                       oob = scales::squish, alpha = 0, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(xmin, xmax), ylim = c(ymin, ymax), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent", colour = NA),
        axis.line = element_blank(),    
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3, colour = NA))
ggsave("plots/map_velo_BB_vector.png", map_BB_vector, bg = "transparent",
       height = 3.55, width = 3.4, units = "in", dpi = 600)
```

Beaufort Sea.

```{r map-BF, warning=FALSE}
# Map limits
xmin <- -28
xmax <- -6
ymin <- -15
ymax <- 9

map_BF <- velo_HD %>%
  ggplot() +
  # Plot interpolated current velocity
  geom_contour_fill(aes(x = x, y = y, z = velocity), binwidth = 0.5, na.fill = T) +
  # Cover areas that are shallower than our data
  geom_tile(data = subset(velo_HD, is.na(vxo) == T), aes(x = x, y = y), fill = "white", col = "white") +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "grey30") +
  # # Current vectors
  # geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), skip = 1, size = 0.15, arrow.angle = 30,
  #            arrow.type = "open", arrow.length = 0.5) +
  # Arctic circle
  geom_path(data = arctic_circle, aes(x = x, y = y), col = "white", lty = 2, size = 0.3) +
  # Plot North Pole
  geom_point(aes(x = 0, y = 0), shape = 4, size = 0.7, stroke = 0.3, col = "white") +
  # GAM locations
  # geom_tile(data = SA_new, aes(x = x, y = y), fill = NA, colour = "red", size = 9) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.8, 
                       oob = scales::squish, alpha = 0.7, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(xmin, xmax), ylim = c(ymin, ymax), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        axis.line = element_blank(),
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3))
ggsave("plots/map_velo_BF.png", map_BF, height = 3.15, width = 3, units = "in", dpi = 600)

map_BF_vector <- velo_HD %>%
 ggplot() +
  # Coast
  geom_polygon(data = coast, aes(x = x, y = y, group = group), fill = "transparent") +
  # Current vectors
  geom_arrow(aes(x = x, y = y, dx = vxo, dy = vyo, mag = velocity), skip = 1, size = 0.2, arrow.angle = 30,
             arrow.type = "open", arrow.length = 0.5) +
  # Scales
  scale_fill_cmocean("", name = "delta", limits = c(0, 5), end = 0.8, 
                       oob = scales::squish, alpha = 0, breaks = seq(0, 10, 0.5)) +
  coord_fixed(xlim = c(xmin, xmax), ylim = c(ymin, ymax), expand = F) +
  guides(mag = "none", alpha = "none", fill = guide_colorsteps(show.limits = T)) +
  theme(legend.position = "none",
        panel.grid = element_blank(),
        plot.background = element_rect(fill = "transparent", colour = NA),
        panel.background = element_rect(fill = "transparent", colour = NA),
        axis.line = element_blank(),    
        axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.border = element_rect(fill = NA, size = 0.3, colour = NA))
ggsave("plots/map_velo_BF_vector.png", map_BF_vector, bg = "transparent",
       height = 3.15, width = 3, units = "in", dpi = 600)
```
